My title

### 2024-11-01 paper/paper.qmd
first draft for model/result/methodology analysis

library(tidyverse)
```

@@

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

@@

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

@@

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

@@

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

# Appendix {#sec-appendix}

# Pollster Methodology Analysis {#sec-appendix-pollster}

@@

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

@@

plot(first_model, "trace")

plot(first_model, "rhat")
```

>>>>

library(tidyverse)
library(here)
library(kableExtra)
library(broom.mixed)
library(knitr)
library(rstanarm)
library(modelsummary)

#### Read in necessary data and model ####
cleaned_president_polls <-read_csv(file = here("data/02-analysis_data/cleaned_president_polls.csv"), show_col_types = FALSE)
cleaned_data_survey <- read_csv(file = here("data/02-analysis_data/national_polling.csv"), show_col_types = FALSE)
state_polling_data <- read_csv(file = here("data/02-analysis_data/state_polling_data.csv"), show_col_types = FALSE)
state_analysis <- read_csv(file = here("data/02-analysis_data/state_analysis.csv"), show_col_types = FALSE)

harris_model <- readRDS(file = here("models/vote_for_Harris_model.rds"))
prediction <- read_csv(file = here("data/02-analysis_data/popular_vote_predictions.csv"), show_col_types = FALSE)
```

@@

We have used a poll-of-polls of the upcoming US presidential election dataset obtained from FiveThirtyEight [@fivethirtyeight_polls]. Data was collected and analyzed using R statistical programming software [@citeR], with additional packages like tidyverse [@tidyverse], rstanarm [@rstanarm] knitr [@citeKnitr], here [@here], and many others for support.

@@

```

@@

```

@@

# Model
For our analysis, we employ a Bayesian generalized linear model (GLM) to forecast the popular vote percentage for Kamala Harris. This approach allows us to capture polling characteristics and account for known variations between pollsters, population types, and the recency of polling data. By incorporating these factors, we aim to model a nuanced estimate of Harris's projected popular vote.

The first step in our process involved selecting a reliable dataset for model development. Here, we utilized high-quality national polling data gathered after Harris’s campaign announcement. We filtered the dataset to include polls with a numeric grade of 2.7 or above, ensuring data quality, and focused only on polls conducted after July 21, 2024, the date of Harris’s declaration of candidacy. This filter helps limit bias from older polls that may not reflect the current voter sentiment.

The GLM is specified as follows:
\begin{equation}
Y_i = \beta_{0} + \beta_{1}x_{pollster_i} + \beta_{2}x_{population_i} + \beta_{3}x_{recentpoll_i} + \epsilon_{i}
\label{eq:logit}
\end{equation}

In equation \ref{eq:logit}, each $\beta$ represents a coefficient determined through regression analysis. The variables chosen for this project are pollster, population type, and recency of the poll. Each predictor variable was carefully selected based on its significance in polling analysis and its correlation with voting trends. The identity of the polling organization is critical to our model, as each pollster may exhibit unique biases. Including Pollster as a fixed effect allows us to account for these variations without introducing unnecessary complexity. Population Type distinguishes among different groups surveyed (e.g., voters, likely voters) and helps in capturing generalizability. Categorizing polls as either recent or not ensures that more recent polls, which are better predictors of voting behavior closer to the election, receive appropriate emphasis. $Y_i$ denotes the predicted popular vote percentage for Kamala Harris in the $i$-th poll. $\epsilon_i$ is the Gaussian-distributed error term, accounting for residual variation in the model.

The selection of predictors in this model is based on observable patterns in the data, such as the consistent tendencies of certain pollsters and the impact of population type on polling outcomes. To enhance the model's robustness, Bayesian priors were applied, introducing regularization and incorporating plausible ranges grounded in previous election data and polling analysis. For the coefficient priors $\beta$, a normal distribution with a mean of 0 and a scale of 2.5 (autoscaled) was chosen to provide flexibility while mitigating overfitting. Similarly, the intercept uses a normal prior with a mean of 0 and scale of 2.5 to stabilize model estimates. For the error term (sigma), an exponential prior with a rate of 1 was selected to constrain the residuals, aligning with Gaussian assumptions. These priors offer a balance between model flexibility and constraint, informed by established trends in polling data and comparable electoral forecasts.

The model was implemented in R [@citeR] using the `rstanarm` package, which offers an accessible interface for Bayesian generalized linear models (GLMs), allowing specification of priors and customization of model parameters. Once the logistic regression model for predicting the popular vote is developed, we apply the `predict()` function in R [@citeR] to generate popular vote percentage predictions for each candidate, using filtered national data. These predictions are then added to the dataset as a new variable (`predicted_pct`) for further analysis. The results are saved in a CSV file (`popular_vote_predictions.csv`), enabling an in-depth examination of potential outcomes in the popular vote. This approach facilitates understanding of vote distributions at a national level and provides a foundation for forecasting electoral outcomes based on demographic and polling data.

# Results
To assess model reliability, we examined several key diagnostics. Convergence metrics, such as Rhat values, were close to 1 for all parameters, indicating strong convergence. Additionally, the effective sample size (n_eff) was high across parameters, suggesting low autocorrelation and contributing to model stability. For validation, we conducted out-of-sample testing and calculated the Root Mean Square Error (RMSE) to assess predictive accuracy. This test data, which includes polls not used in training, provides an unbiased estimate of model performance.

The model operates under the assumption that residuals follow a Gaussian distribution, though this may not entirely capture extreme polling variances. Additionally, pollster effects are treated as fixed, which, while limiting the model’s ability to reflect dynamic shifts in polling methodologies, simplifies the model and reduces the risk of overfitting. Alternative specifications, including a hierarchical model with random intercepts for pollsters, were considered. However, the added complexity of this approach did not significantly improve predictive accuracy. The chosen GLM specification thus offered a balanced approach, optimizing both interpretability and performance, making it the preferred model for our study.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-model-coefficients
#| tbl-cap: Coefficients from the GLM Model

# Tidy and round the model summary data
coefficients <- broom.mixed::tidy(harris_model, conf.int = TRUE) %>%
  mutate(across(c(estimate, std.error, conf.low, conf.high), round, 2))

# Generate the LaTeX table with 2 decimal places
kable(coefficients, format = "latex", booktabs = TRUE, align = "c") %>% 
  kable_styling(latex_options = "scale_down", font_size = 7)
```

Table @tbl-model-coefficients presents the estimated coefficients for the predictors in our GLM model. These coefficients fit into the GLM equation, allowing us to interpret the impact of each predictor on Harris’s predicted vote percentage. Positive values indicate a higher predicted percentage, while negative values indicate a decrease. Key predictors, such as specific pollsters and population types, show distinct effects on the forecasted outcomes. This table is produced using kable from knitr for enhanced readability.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-coefficient-estimates
#| fig-cap: Coefficient Estimates for Predictors
coefficients %>%
  ggplot(aes(estimate, term)) + 
  geom_point() + 
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) + 
  labs(title = "Coefficient Estimates", x = "Estimate", y = "Predictor") +
  theme(axis.text.y = element_text(size = 7))
```

Figure @fig-coefficient-estimates visually represents the model coefficients, with error bars indicating the confidence interval for each estimate. Positive coefficients suggest that specific pollsters or demographic characteristics increase Harris’s predicted vote share. For example, the 'CES / YouGov' pollster shows a significant positive impact, while 'Siena/NYT' has a slight negative effect. These error bars help contextualize the reliability of each predictor.

@@

# YouGov Methodology Analysis {#sec-appendix-pollster}

YouGov employs a panel-based methodology for conducting online surveys, recruiting participants who voluntarily join their platform. Their surveys span topics like politics, policy opinions, and voter behavior. In the context of election polling, YouGov applies Multilevel Regression and Post-Stratification (MRP) models to predict outcomes by blending respondent data with external data like voter files [@yougov_mrp_2024].

The target population for YouGov’s surveys typically consists of U.S. adults, with specific samples drawn for each survey depending on the topic. YouGov’s sample frame consists of individuals who have opted into their online panel. To ensure representativeness, YouGov sets demographic quotas, stratifying by variables such as age, gender, and political affiliation. Additionally, YouGov monitors and adjusts their sample post-survey using MRP to correct for demographic imbalances and potential sampling bias.

YouGov’s panel is recruited through various online methods, including digital advertisements and direct outreach. This recruitment strategy leads to a convenience sample, meaning participants self-select to join the panel, which can introduce selection bias. Although YouGov attempts to offset this bias using quotas and post-stratification weighting, the sample still primarily consists of individuals comfortable with online surveys. This method may under-represent certain demographics, such as older adults or individuals with limited internet access.

YouGov’s use of non-probability sampling—drawing respondents from a convenience sample—provides cost efficiency and quick data collection. However, the trade-off lies in the introduction of potential biases due to self-selection. While YouGov’s MRP model corrects some biases by leveraging population-level data for small samples, the non-random nature of the sample may still skew results. For instance, respondents who engage in online surveys may systematically differ from those who avoid them, such as individuals with lower levels of social trust.

Non-response bias is a significant concern in YouGov’s methodology. Individuals who opt out of surveys may systematically differ from those who participate. YouGov employs several strategies to mitigate this issue, including **post-survey weighting**, which adjusts the survey data to reflect population characteristics, and **imputation techniques**, which estimate missing responses based on patterns in the available data. Despite these efforts, non-response bias can still affect the accuracy of predictions, particularly when certain demographic groups (e.g., less-educated or low-trust individuals) are less likely to participate.

YouGov places emphasis on clear and concise survey design, reducing participant fatigue and improving data quality. Their surveys generally avoid overly complex or leading questions, which could skew results. However, question wording remains a critical area, as even subtle changes in phrasing can influence responses. For instance, YouGov’s research shows how question wording can significantly affect voter perception and reported support for policies. Additionally, longer surveys can lead to increased dropout rates, potentially introducing non-response bias. To combat this, YouGov often incorporates survey logic, dynamically adjusting questions based on prior responses to keep participants engaged.

YouGov’s panel-based methodology, combined with advanced statistical techniques like MRP, offers a robust approach to polling. Their methods effectively address many challenges associated with non-random sampling and non-response bias. However, limitations persist, particularly regarding the biases inherent in self-selection and question framing. Overall, while YouGov’s methodology is well-suited to large-scale surveys, particularly in politics, careful consideration of their sampling and non-response strategies is necessary to fully interpret their findings.

@@

#pp_check(first_model) +
#  theme_classic() +
 # theme(legend.position = "bottom")

#posterior_vs_prior(first_model) +
#  theme_minimal() +
#  scale_color_brewer(palette = "Set1") +
 # theme(legend.position = "bottom") +
 # coord_flip()
```

@@

#plot(first_model, "trace")

#plot(first_model, "rhat")
```
